---
title: "Clustring task"
author: "Yunfang Wu"
date: "2022-08-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r 'load_packages', warning = FALSE, message = FALSE}
# --- load libraries
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(sf)
library(furrr)
library(lubridate)
library(rnaturalearth)
library(gganimate)
#library(bomrang)
library(spotoroo)
library(ggmap)
library(gridExtra)
```
# Import hotspot data

Firstly, dowload hotspot data from ftp://ftp.ptree.jaxa.jp/pub/himawari/L3/WLF/bet, 
Hotspot data from 2021 to 2022 has been dowloaded,
to test accuracy of clustring against previous report, 
 
## First, read all monthly data from 201710 to 201803 into one data frame
```{r}
#setwd('/Users/vickywu/Desktop/master_project/data/Himawari-8')
#df <- list.files() %>% 
#  lapply(read_csv) %>% 
 # bind_rows 
```


```{r 'define_read_hotspot file'}
#read_hotspot <- function(directory = "data/Himawari-8", from = '20190501', to = '20190701'){
  ### include from, exclude to
  
 # fls = list.files(directory, full.names = TRUE)
 # fls_short = list.files("data/Himawari-8", full.names = FALSE)
#  fls_filter = (fls_short > paste('H08', from, sep = '_')) & (fls_short < paste('H08', to, sep = '_'))
 # fls = fls[fls_filter]
 # all = map(fls, read_csv, quote="'")
 # d = bind_rows(all)
 # return(d)
  
#}
```
```{r 'read_hotspots if in file', warning = FALSE, message = FALSE}
#hotspots = read_hotspot(directory = "data/Himawari-8", from = '20171001', to = '20180300')
```

```{r 'read hotspot data form csv'}
# only
recent_hotspots <- read_csv(here("./data/2019-2020hotpots.csv"))
```
```{r 'read hotspot data form csv'}
# only
all_hotspots <- read_csv(here("./data/all_time_hotpots.csv"))
```

```{r}
recent_hotspots %>% 
  st_as_sf(., coords = c("longitude","latitude"), remove=FALSE, crs = 4326)->recent_hotspots_sf

```

```{r}
all_hotspots %>% 
  st_as_sf(., coords = c("longitude","latitude"), remove=FALSE, crs = 4326)->all_hotspots_sf
```


```{r}
# get victoria map
ne_states(country = "Australia", returnclass ="sf") %>% 
  filter(gns_name=="Victoria") ->vic_map
```

```{r}
# get all hotspots in Australia
ne_states(country = "Australia", returnclass ="sf") -> au_map
```

```{r}
all_hotspots_au <- all_hotspots_sf[au_map, ]
```


# verify the polygon and point data have the same crs 
This step make sure we sf points object has the same crs as the polygon object

```{r}
st_crs(vic_map)
```


```{r}
all_hotspots_vic <- all_hotspots_sf[vic_map, ]
```

```{r}
recent_hotspots_au<- recent_hotspots_sf[au_map,]
```



```{r 'get_au_hotspots'}
# apply australia coordinates 
#hotspots_au = hotspots %>%
#  filter(between(lon, 112, 155)) %>% 
#  filter(between(lat, -44, -10))
```

```{r 'fire_power_filter'}
#firepower_threshold = 100 #threshold for firepower
#hotspots_vic <- hotspots_vic %>%
#  filter(power > firepower_threshold)
```




```{r}
samples = dplyr::select(all_hotspots_vic, `datetime`, `longitude`, `latitude`, `power`)
p = ggplot(samples) +
  geom_sf(data = vic_map) +
  geom_point(aes(group = `datetime`, x = longitude, y = latitude, size = power), color = 'red', alpha = 0.2) + 
  transition_time(`datetime`) +
  exit_fade() +
  enter_fade() +
  labs(title = 'Time: {frame_time}') +
  theme_bw()

```

# visualize fire hotspot with time
```{r}
animate(p)
```
#apply clustering to au recent hotspot data

```{r}
#select 
recent_hotspots_au_df<- recent_hotspots_au %>% 
  dplyr::select(`datetime`, longitude, latitude) %>% 
  as.data.frame()
```

```{r}
all_vic_hotspots_df<- all_hotspots_vic %>% 
  dplyr::select(`datetime`, longitude, latitude) %>% 
  as.data.frame()
```


```{r}
all_result <- hotspot_cluster(all_vic_hotspots_df,
                          lon = "longitude",
                          lat = "latitude",
                          obsTime = 'datetime',
                          activeTime = 24,
                          adjDist = 3000,
                          minPts = 4,
                          minTime = 3,
                          ignitionCenter = "mean",
                          timeUnit = "h",
                          timeStep = 1)
```


```{r}
result <- hotspot_cluster(all_vic_hotspots_df,
                          lon = "longitude",
                          lat = "latitude",
                          obsTime = 'datetime',
                          activeTime = 24,
                          adjDist = 3000,
                          minPts = 4,
                          minTime = 3,
                          ignitionCenter = "mean",
                          timeUnit = "h",
                          timeStep = 1)
```
```{r}
result
```

```{r}
summary(result)
```
### Plotting 2D density estimate

1. extract ignition information and store it into a dataframe

```{r}
ignitions<-result[["ignition"]]
```



## using ggmap package 

```{r}
# create other types of plots with the ggmap package
densityplot_ignitions <- qmplot(x=lon, y=lat, 
                             data = ignitions, 
                             geom = "blank",
                             maptype = "toner-background", 
                             darken = .2, 
                             legend = "topright") + stat_density_2d(aes(fill = ..level..), 
                  geom = "polygon", 
                  alpha = .5,
                  color = NA) + scale_fill_gradient2(low = "blue", 
                       mid = "green", 
                       high = "red")
```
```{r}
plot(densityplot_ignitions)
```
## kernal density estimation



```{r}
library(ggthemes)
```









