---
title: "model"
author: "Yunfang Wu"
date: "2022-10-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(tidyverse)
library(here)
library(rpart)
library(rpart.plot)
library(randomForest)
library(knitr)
```

```{r include=FALSE}
#knitr::knit_child("sample_density_estimation.Rmd")
```


1. Read in data

```{r}
sample_hotspots <-read_csv(here::here("data/sample_estimate.csv"))
```


```{r}


library(fastDummies)
sample_hotspots<- dummy_cols(sample_hotspots,select_columns="FOR_TYPE")
```

remove where dummy is empty
```{r}
names(sample_hotspots) <- gsub(" ", "_", names(sample_hotspots))
names(sample_hotspots) <- gsub("\\?", "_", names(sample_hotspots))
sample_hotspots


sample_hotspots<-sample_hotspots %>% 
  select(-c(code,long,lat,new_cause,wod,day,FOR_TYPE)) %>% 
  #covert all eucalypt stpe to one row
  mutate(FOR_TYPE_Eucalypt = FOR_TYPE_Eucalypt_Low_Open+
                                                FOR_TYPE_Eucalypt_Low_Woodland+
                                                FOR_TYPE_Eucalypt_Mallee_Woodland+
                                                FOR_TYPE_Eucalypt_Medium_Closed+
                                                FOR_TYPE_Eucalypt_Medium_Open+
                                                FOR_TYPE_Eucalypt_Medium_Woodland+
                                                FOR_TYPE_Eucalypt_Tall_Closed+
                                               FOR_TYPE_Eucalypt_Tall_Open+
                                               FOR_TYPE_Eucalypt_Tall_Woodland)%>% 
  select(-c(FOR_TYPE_Eucalypt_Low_Open,
                                                FOR_TYPE_Eucalypt_Low_Woodland,
                                                FOR_TYPE_Eucalypt_Mallee_Woodland,
                                                FOR_TYPE_Eucalypt_Medium_Closed,
                                                FOR_TYPE_Eucalypt_Medium_Open,
                                                FOR_TYPE_Eucalypt_Medium_Woodland,
                                                FOR_TYPE_Eucalypt_Tall_Closed,
                                               FOR_TYPE_Eucalypt_Tall_Open,
                                               FOR_TYPE_Eucalypt_Tall_Woodland))
```




Data description:
FOR_TYPE: Forest type. Eg. Acacia, Callitris, Casuarina, etc.
COVER: Forest crown cover
HEIGHT: Forest height class
arf360: Average rainfall in the past 360 days in mm
ase90: Average global solar exposure in the past 90 days - MJ/m^2
ase180: Average global solar exposure in the past 180 days - MJ/m^2
ase720: Average global solar exposure in the past 720 days - MJ/m^2
amaxt90: Average maximum temperature in the past 90 days - Celsius degree
amaxt180: Average maximum temperature in the past 180 days - Celsius degree
amaxt720: Average maximum temperature in the past 720 days - Celsius degree
amint180: Average minimum temperature in the past 180 days - Celsius degree
ws: Average wind speed on that day - m/s
aws_m12: Average wind speed in last 12 months - m/s
aws_m24: Average wind speed in last 24 months - m/s
log_dist_cfa: Natural logarithm of the distance to the nearest CFA station - m
log_dist_camp: Natural logarithm of the distance to the nearest recreation site - m
log_dist_road: Natural logarithm of the distance to the nearest road - m





2. Now split the data to training and testing sets, have the training set be 70% of the data, and the test set be 30% of the data.

```{r}
set.seed(123)
model_df <- sample_hotspots%>%
  mutate(test_or_train = 
           sample(c("test","train"),n(),replace = TRUE, prob = c(.3,.7))) 

model_train <- model_df %>%
  filter(test_or_train == "train") %>% 
  select(-c("test_or_train"))

model_test <- model_df %>%
  filter(test_or_train == "test") %>% 
  select(-c("test_or_train"))

```

3.Now use all the variables described above to create a regression tree with the price as an outcome using the training data only.






```{r}
set.seed(222)
reg_tree <- rpart(density_result~. , data = model_train, 
  minbucket = 1)
```

4. Create a visual representation of the tree 
```{r}
rpart.plot(reg_tree)
```

6. Use the tree to predict for the test part of the data

```{r}
set.seed(222)
predictions_tree <- predict(reg_tree, model_test)
```


7. Calculate the root mean square error

```{r}
model_test <- model_test %>%
  mutate(density_tree_preds = predictions_tree)

model_test %>%
  summarise(MSE = (mean((density_result - density_tree_preds)^2)))

```
8. Compare this to the RMSE for a linear model with just the predictors used in the tree.
```{r}
reg_lm <- lm(density_result~., data = model_train)

model_test <-model_test %>%
  mutate(density_lm_preds = predict(reg_lm, model_test))

model_test %>%
  summarise(MSE = mean((density_result - density_lm_preds)^2))

```

Now fit a random forest model.
```{r}
write.csv(sample_hotspots,"model_df.csv", row.names = FALSE)
```

```{r}
sample_hotspots <-read_csv(here::here("model_df.csv"))
set.seed(456)
RFM = randomForest(density_result~., data = sample_hotspots,importance=TRUE)
```

```{r}
print(RFM)
```
now change the default parameter, fit 1000 trees

```{r}
set.seed(678)
RFM2 = randomForest(density_result~., ntree =1000, mtry = 25, importance = TRUE,data = sample_hotspots)

```
compare result

```{r}
RFM2
```

results are similar


```{r}
importance(RFM,type=1)
```
```{r}
varImpPlot(RFM, scale=TRUE,n.var=13,main="Variable Importance Plot")
```









Data description:
FOR_TYPE: Forest type. Eg. Acacia, Callitris, Casuarina, etc.
COVER: Forest crown cover
HEIGHT: Forest height class
arf360: Average rainfall in the past 360 days in mm
ase90: Average global solar exposure in the past 90 days - MJ/m^2
ase180: Average global solar exposure in the past 180 days - MJ/m^2
ase720: Average global solar exposure in the past 720 days - MJ/m^2
amaxt90: Average maximum temperature in the past 90 days - Celsius degree
amaxt180: Average maximum temperature in the past 180 days - Celsius degree
amaxt720: Average maximum temperature in the past 720 days - Celsius degree
amint180: Average minimum temperature in the past 180 days - Celsius degree
ws: Average wind speed on that day - m/s
aws_m12: Average wind speed in last 12 months - m/s
aws_m24: Average wind speed in last 24 months - m/s
log_dist_cfa: Natural logarithm of the distance to the nearest CFA station - m
log_dist_camp: Natural logarithm of the distance to the nearest recreation site - m
log_dist_road: Natural logarithm of the distance to the nearest road - m





```{r}
model_test <- model_test %>%
  mutate(density_forest_preds = predictions_forest)

model_test %>%
  summarise(MSE = mean((density_result - density_forest_preds)^2))
```
We can see that the random forest model has the lowest RMSE.


```{r}
sample_hotspots %>% 
  tibble::rowid_to_column(., "ID") %>% 
  select(c(dist_camp,dist_cfa,amint720,amint720,arf720,amaxt60,
           aws_m24,COVER,HEIGHT,FOR_TYPE_Eucalypt,density_result,ID)) %>% 
  pivot_longer(cols = c(amint720,amint720,arf720,amaxt60,aws_m24
                        ,dist_camp,dist_cfa,COVER,HEIGHT,FOR_TYPE_Eucalypt), 
               names_to="variable",
               values_to = "values") %>% 
  group_by(ID, variable) %>% 
  ggplot(.,aes(x=values,y=density_result)) +
  geom_point() +
  facet_wrap(~variable,scales = "free")
```
Average wind speed in last 24 months

```{r}
example<-sample_hotspots[1,]
```

```{r}
lam=matrix(c())
```


```{r}
sample_hotspots
```

