---
title: "Exploring bushfire risk across Victoria with open data sources and open-source software"
format: revealjs
editor: visual
---

## Motivation

Victoria, Australia is one of the most fire-prone regions in the world - given its fire-conducive weather and fuel conditions.

```{r}
#| echo: false
#| fig-cap: "Historical fire hotspots in the state of Victoria, Australia"

#samples = dplyr::select(all_hotspots_vic, `datetime`, `longitude`, `latitude`, `power`)
#p = ggplot(samples) +
#  geom_sf(data = vic_map) +
#  geom_point(aes(group = `datetime`, x = longitude, y = latitude, size = power), color = 'red', alpha = 0.2) + 
 # transition_time(`datetime`) +
 # exit_fade() +
 # enter_fade() +
 # labs(title = 'Time: {frame_time}') +
 # theme_bw()

#animate(p)
```

## Research Questions

-   What part of Victoria has higher bush fire risk?
-   What factors are the more important for predicting the risk accurately
-   How change in factors affect the bush fire risk of a particular location?

## Methodology

### What part of Victoria has higher bush fire risk?

-   Kernal density estimation

    a [non-parametric](https://en.wikipedia.org/wiki/Non-parametric_statistics "Non-parametric statistics") method to [estimate](https://en.wikipedia.org/wiki/Estimation "Estimation") the [probability density function](https://en.wikipedia.org/wiki/Probability_density_function "Probability density function") of a [random variable](https://en.wikipedia.org/wiki/Random_variable "Random variable") based on [*kernels*](https://en.wikipedia.org/wiki/Kernel_(statistics) "Kernel (statistics)") as [weights](https://en.wikipedia.org/wiki/Weight_function "Weight function").

```{r}
# Data
set.seed(14012021)
data <- rnorm(200, mean = 4)

# Kernel density estimation
d <- density(data)

# Kernel density plot
plot(d, lwd = 2, main = "Default kernel density plot")
```

## Methodology

### What part of Victoria has higher bush fire risk?

```{r}
library(MASS)
attach(geyser)
plot(duration, waiting, xlim = c(0.5,6), ylim = c(40,100))
f1 <- kde2d(duration, waiting, n = 50, lims = c(0.5, 6, 40, 100))
image(f1, zlim = c(0, 0.05))
f2 <- kde2d(duration, waiting, n = 50, lims = c(0.5, 6, 40, 100),
            h = c(width.SJ(duration), width.SJ(waiting)) )
image(f2, zlim = c(0, 0.05))
persp(f2, phi = 30, theta = 20, d = 5)

plot(duration[-272], duration[-1], xlim = c(0.5, 6),
     ylim = c(1, 6),xlab = "previous duration", ylab = "duration")
f1 <- kde2d(duration[-272], duration[-1],
            h = rep(1.5, 2), n = 50, lims = c(0.5, 6, 0.5, 6))
contour(f1, xlab = "previous duration",
        ylab = "duration", levels  =  c(0.05, 0.1, 0.2, 0.4) )
f1 <- kde2d(duration[-272], duration[-1],
            h = rep(0.6, 2), n = 50, lims = c(0.5, 6, 0.5, 6))
contour(f1, xlab = "previous duration",
        ylab = "duration", levels  =  c(0.05, 0.1, 0.2, 0.4) )
f1 <- kde2d(duration[-272], duration[-1],
            h = rep(0.4, 2), n = 50, lims = c(0.5, 6, 0.5, 6))
contour(f1, xlab = "previous duration",
        ylab = "duration", levels  =  c(0.05, 0.1, 0.2, 0.4) )
detach("geyser")
```

-   Bias of density estimation of edges (or frontiers) of regions

    -   Broader bias is when

    -   Ripley's correction

```{r}
#child = "sample_density_estimation.Rmd"
    #| echo: false
    #| fig-cap: "Desity estimation with correction of broader bias"


  #  print(p_igni)
```
