---
title: "model"
author: "Yunfang Wu"
date: "2022-10-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(tidyverse)
library(here)
library(rpart)
library(rpart.plot)
library(randomForest)
library(knitr)
```

```{r include=FALSE}
knitr::knit_child("sample_density_estimation.Rmd")
```


1. Read in data

```{r}
sample_hotspots <-read_csv(here::here("data/sample_estimate.csv"))
```


```{r}
sample_hotspots<-sample_hotspots %>% 
  select(-c(code,long,lat,new_cause,wod,day))
```
```{r}
colnames(sample_hotspots)
```



Data description:
FOR_TYPE: Forest type. Eg. Acacia, Callitris, Casuarina, etc.
COVER: Forest crown cover
HEIGHT: Forest height class
arf360: Average rainfall in the past 360 days in mm
ase90: Average global solar exposure in the past 90 days - MJ/m^2
ase180: Average global solar exposure in the past 180 days - MJ/m^2
ase720: Average global solar exposure in the past 720 days - MJ/m^2
amaxt90: Average maximum temperature in the past 90 days - Celsius degree
amaxt180: Average maximum temperature in the past 180 days - Celsius degree
amaxt720: Average maximum temperature in the past 720 days - Celsius degree
amint180: Average minimum temperature in the past 180 days - Celsius degree
ws: Average wind speed on that day - m/s
aws_m12: Average wind speed in last 12 months - m/s
aws_m24: Average wind speed in last 24 months - m/s
log_dist_cfa: Natural logarithm of the distance to the nearest CFA station - m
log_dist_camp: Natural logarithm of the distance to the nearest recreation site - m
log_dist_road: Natural logarithm of the distance to the nearest road - m





2. Now split the data to training and testing sets, have the training set be 70% of the data, and the test set be 30% of the data.

```{r}
set.seed(123)
model_df <- sample_hotspots%>%
  mutate(test_or_train = 
           sample(c("test","train"),n(),replace = TRUE, prob = c(.3,.7)))

model_train <- model_df %>%
  filter(test_or_train == "train")

model_test <- model_df %>%
  filter(test_or_train == "test")

```

3.Now use all the variables described above to create a regression tree with the price as an outcome using the training data only.

```{r}
set.seed(123)
reg_tree <- rpart(density_result~FOR_TYPE+COVER+HEIGHT+rf+arf7+arf14+arf28+arf60+arf90+arf180       +arf360+arf720+se+ase7+ase14+ase28+ase60+ase90+ase180+ase360+ase720+maxt+amaxt7+amaxt14+amaxt28+amaxt60+amaxt90+amaxt180+amaxt360+amaxt720+mint+amint7+amint14+amint28+amint60+amint90+amint180+month , data = model_train)
```

4. Create a visual representation of the tree 
```{r}
rpart.plot(reg_tree)
```

6. Use the tree to predict for the test part of the data

```{r}
set.seed(123)
predictions_tree <- predict(reg_tree, model_test)
```


7. Calculate the root mean square error

```{r}
model_test <- model_test %>%
  mutate(density_tree_preds = predictions_tree)

model_test %>%
  summarise(RMSE = sqrt(mean((density_result - density_tree_preds)^2)))

```
8. Compare this to the RMSE for a linear model with just the predictors used in the tree.
```{r}
reg_lm <- lm(density_result~FOR_TYPE+COVER+HEIGHT+rf+arf7+arf14+arf28+arf60+arf90+arf180       +arf360+arf720+se+ase7+ase14+ase28+ase60+ase90+ase180+ase360+ase720+maxt+amaxt7+amaxt14+amaxt28+amaxt60+amaxt90+amaxt180+amaxt360+amaxt720+mint+amint7+amint14+amint28+amint60+amint90+amint180+month, data = model_train)

model_test <-model_test %>%
  mutate(density_lm_preds = predict(reg_lm, model_test))

model_test %>%
  summarise(RMSE = sqrt(mean((density_result - density_lm_preds)^2)))

```

Now fit a random forest model.

```{r}
set.seed(456)
RFM = randomForest(density_result~FOR_TYPE+COVER+HEIGHT+rf+arf7+arf14+arf28+arf60+arf90+arf180       +arf360+arf720+se+ase7+ase14+ase28+ase60+ase90+ase180+ase360+ase720+maxt+amaxt7+amaxt14+amaxt28+amaxt60+amaxt90+amaxt180+amaxt360+amaxt720+mint+amint7+amint14+amint28+amint60+amint90+amint180+month, data = model_train)
```

```{r}
print(RFM)
```

0.002501067	

6. Use the random forest model to predict for the test part of the data

```{r}
predictions_forest <- predict(RFM, model_test)
```

```{r}
imp <- as.data.frame(sort(importance(RFM)[,1],decreasing = TRUE),optional = T)
names(imp) <- "% Inc MSE"
imp
```
Data description:
FOR_TYPE: Forest type. Eg. Acacia, Callitris, Casuarina, etc.
COVER: Forest crown cover
HEIGHT: Forest height class
arf360: Average rainfall in the past 360 days in mm
ase90: Average global solar exposure in the past 90 days - MJ/m^2
ase180: Average global solar exposure in the past 180 days - MJ/m^2
ase720: Average global solar exposure in the past 720 days - MJ/m^2
amaxt90: Average maximum temperature in the past 90 days - Celsius degree
amaxt180: Average maximum temperature in the past 180 days - Celsius degree
amaxt720: Average maximum temperature in the past 720 days - Celsius degree
amint180: Average minimum temperature in the past 180 days - Celsius degree
ws: Average wind speed on that day - m/s
aws_m12: Average wind speed in last 12 months - m/s
aws_m24: Average wind speed in last 24 months - m/s
log_dist_cfa: Natural logarithm of the distance to the nearest CFA station - m
log_dist_camp: Natural logarithm of the distance to the nearest recreation site - m
log_dist_road: Natural logarithm of the distance to the nearest road - m





```{r}
model_test <- model_test %>%
  mutate(density_forest_preds = predictions_forest)

model_test %>%
  summarise(MSE = mean((density_result - density_forest_preds)^2))
```
We can see that the random forest model has the lowerst RMSE.

pick one hotspot, use 
-predicted vs obs,
-residual plot
-obs, vs 3 most important var

